Numerical error propagation in the HCP structural pre-processing pipelines

### 2018-10-01 paper.tex

\usepackage{pdfcomment}
\usepackage[margin=1in]{geometry}
\usepackage{xcolor}
\usepackage{graphicx}
\usepackage{algorithm}
\usepackage{caption}
\usepackage{listings}
\usepackage{verbatim}

@@

We randomly selected 10 subjects from the HCP data release
S500. \todo{ list the subjects.}

@@

@@

@@

We reconstruct the tree of processes starting from the first process
created by the pipeline and identifying its child processes as the ones
that were created through \texttt{clone()} or \texttt{fork()}. We
create a process graph from this tree by adding edges corresponding to
file dependencies between processes. A file dependency is defined
between processes A and B if a file written by A is read by
B. Figure~\ref{fig:graph-example} shows an example of a process graph
constructed at this stage.

@@

bet ${input_image} ${bet_output} > bet_temp.out
echo "Voxels / Volume in brain mask:"
fslstats ${bet_output} -V
fslmaths ${bet_output} ${bet_output_binarized}
echo "Voxels / Volum in binarized brain mask:"
\rm bet_temp.out
\end{verbatim}
}
\end{minipage}
\captionof{figure}{Process graph constructed from an example brain extraction
    script. Every node in the graph is labeled using (1) a process id
    created by our reconstruction, (2) the name of the executable run
    by the process. Processes that read or write temporary files are
    represented with squares; other ones with circles. Plain edges
    represent the process tree (\texttt{fork()} or \texttt{clone()}
    system calls). Dashed edges represent file dependencies: temporary
    files are in yellow and result files are in green.} \label{myfig1}
\medskip

@@

\paragraph{} In the next step, we classify the processes into three 
categories based on the process graph and the error matrix file 
mentioned in the previous Section:
\begin{enumerate}
\item Processes 
that read files that do not have errors and write files that do not 
have errors are \emph{transparent}.
\item Processes that read files 
that do not have any error but write files that have errors 
\emph{create} errors in the pipeline.
\item Processes that read files 
that have errors and write files that do not have errors\emph{remove} 
errors from the pipeline.
\end{enumerate}

The classification of processes that read or write temporary files is 
uncertain when these files are deleted during the execution. To address 
this issue, we replace every process P that writes temporary files with 
a modification of P that first calls P and then backs up all its output 
files to a read-only directory. This replacement is done by modifying 
the Unix PATH variable to point to a directory containing the modified
versions of the processes. This solution does not cover the temporary files
that are removed by P itself; this is not a problem since these files do not
play any role in the subsequent steps of the pipeline, by definition.

Files written by multiple processes also lead to unknown classification 
labels. To address this issue for a file F written by processes in 
\textbf{P} = \{$P_{1}$, \ldots $P_{n}$\}, we (1) check that processes 
in \textbf{P} do not write concurrently to F, (2) we establish an order 
on \textbf{P} based on the creation timestamp of the processes, (3) we 
replace ever process $P_{i}$ in \textbf{P} by a wrapper that first 
calls $P_{i}$ and then backs up F to a read-only directory. Thus, 
multiple versions of F are saved and used in the analysis.

Another challenge is to classify processes that read files that have
errors and write files that also have errors. In such situations, it
is not possible to determine from the pipeline results whether
the process created errors, removed some errors, or was
transparent. We label such processes \emph{unknown}. To address this issue, we developed an iterative approach that consists of the following steps:
\begin{enumerate}
  \item Run the pipeline in conditions 1 and 2; classify the
    processes as \emph{transparent}, \emph{create errors},
    \emph{remove errors} or \emph{unknown}.
  \item If there are \emph{unknown} processes then: replace each process P that create errors by a process Q that
    copies the results produced by process P in condition 1 to the pipeline output in condition 2.
  \item Repeat steps 1 and 2 until there is no \emph{unknown} process left.
\end{enumerate}
The replacement of process P at step 2 is done by replacing P
with a custom script produced by step 1. This custom script
copies the results obtained in condition 1 if it is invoked with the
arguments of a process that created errors, and calls P's original
executable otherwise. The replacement is done through the PATH variable, as before. This algorithm converges to a process graph
without any \emph{unknown} process after a finite number of iterations.

\begin{figure}
  \includegraphics[scale=0.6]{images/iterative_modif}
  \caption{iteration of proposed modification.}
  \label{fig:iterations}
\end{figure}

@@

In a previous study~\cite{Scaria2017}, we showed that
pre-processing pipelines of the Human Connectome
Project~\cite{Glasser2013} were sensitive to operating system
variation (see Figure \ref{fig:1}).
\begin{figure}
%  \includegraphics{brain\_classification}
  \includegraphics[width=0.5\textwidth, height=6cm]{images/brain_classification.png}
  \caption{Tissue classification produced by the HCP pre-processing
    pipelines on subject 105216 (CentOS 6 vs CentOS 7).}
  \label{fig:1}
\end{figure}

@@

Figure \ref{fig:summarized-graph} shows the annotated provenance graph of the
PreFreeSurfer pipeline executed on CentOS6 and CentOS7.  Processes
that created errors are shown in red, processes that removed errors
are in blue, and other processes are in green.  Squares denote
processes for which the classification is uncertain, due to temporary
files that were removed during the execution. Black edges link
sub-processes to their parents while dashed edges denote file
dependencies between processes (green edges: files with no errors; red
edges: files with errors; yellow edges: temporary files).  The
processes that introduce errors in PreFreeSurfer are: linear
registration with “\emph{FLIRT}” (8 occurrences, in ACPC Alignment,
BrainExtraction, DistortionCorrection, AtlasRegistration), non-linear
registration with “\emph{FNIRT}” (3 occurrences, in BrainExtraction
and AltasRegistration), image warping with “\emph{new\_invwarp}” (3
occurrences, in BrainExtraction and AtlasRegistration).  In addition,
errors were observed in image mean and standard-deviation computations
with “\emph{fslstats}” (3 occurrences in BiasFieldCorrection), and in
masked image extrapolation with “\emph{fslmaths}” (1 occurrence, in
BiasFieldCorrection).  Besides, transformation format conversion with
“\emph{convertwarp}” (2 occurrences, in DistortionCorrection) was able
to remove errors.

@@

\begin{figure}
  \includegraphics[width=\linewidth]{images/summarized_graph}
  \caption{An annotated summarized process graph from the PreFreesurfer pipeline.
Each node in the graph is represented in red, blue or green colors and is labeled 
using the name of the executable run by the interesting process. The close up nodes 
reveal each shrinked subgraph in details.}
  \label{fig:summarized-graph}
\end{figure}

\begin{figure}
  \includegraphics[width=\linewidth]{images/graph}
  \caption{A complete process graph from the PreFreesurfer pipeline.
Full-resolution image available at \url{https://drive.google.com/open?id=174yyn8SuVOUcK5aRVw0bagjDanLD0FLt}.}
  \label{fig:complete-graph}
\end{figure}

@@

Discuss the results.

Mention that this was only possible because the unprocessed data was shared in the first place.

DICOM to Nifti conversion was out of scope and may introduce other issues.

@@

Conclude and highlight future work.

The numerical instability in the PreFreesurfer HCP pipeline arises mainly from linear and non-linear registration processes 
implemented in FSL FLIRT and FNIRT. These processes need to be reviewed to understand and correct the cause of instabilities. 
In this correction process, accuracy has to be considered in addition to stability. Our technique is able to 
characterize the stability of a pipeline’s components automatically, but it also suffers from limitations as it cannot deal with: 
(1) results that are not written to disk (values processed in memory); (2) temporary files (3) files that are written by more than one process.

@@

\bibliographystyle{plain}
\bibliography{biblio}

>>>>

\usepackage{pdfcomment}
\usepackage[margin=1in]{geometry}
\usepackage{xcolor}
\usepackage{graphicx}
\usepackage{algorithm}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{listings}
\usepackage{verbatim}
\usepackage{booktabs}
\usepackage[flushleft]{threeparttable}

@@

We randomly selected 10 subjects from the HCP data release
S500 which are described in table~\ref{table:data}. \todo{ list the subjects.}

\begin{table}[H]\hspace*{-1cm}
\centering
\begin{threeparttable}
\caption{An overview of the HCP data (Humman Connectome DB).}

\begin{tabular}{@{}llllll@{}}
\toprule
Subject & Release & Acquisition & Gender & Age      & Subj\_Type   \\ \midrule
103515  & Q1      & Q02         & F      & 26-30    & type1         \\
105216  & Q3      & Q03         & M      & 26-30    & type4         \\
103414  & Q2      & Q02         & F      & 22-25    & type1         \\
125525  & Q1      & Q01         & F      & 31-35    & type1         \\
142828  & Q1      & Q01         & M      & 31-35    & type1         \\
129533  & Q3      & Q04         & F      & 31-35    & type1         \\
103818  & Q1      & Q01         & F      & 31-35    & type3         \\
133928  & Q2      & Q03         & M      & 26-30    & type3         \\
148032  & Q3      & Q03         & F      & 31-35    & type2         \\
139637  & Q2      & Q03         & F      & 31-35    & type1         \\ \bottomrule
\end{tabular}
\begin{tablenotes}
      \small
      \item *Subj\_Type refers to the graph topology type of the subjects,
      we will discuss further on the next sections.
\end{tablenotes}
\end{threeparttable}
\label{table:data}
\end{table}

@@

we focused on HCP structural pre-processing pipelines to:
\begin{itemize}
\item Quantify the effect of OS on HCP pre-processing pipelines.
\item Identify the process(es) in the pipelines that are responsible for such effects.
\end{itemize}

@@

\begin{figure}[H]
\centering
  \includegraphics[scale=0.9]{images/overview.png}
  \caption{An overview of proposed technique.}
  \label{fig:overview}
\end{figure}

Figure~\ref{fig:overview} illustrate an overview of the proposed technique. First, we execute pipelines in a
consistence environments and then the origin of errors in the pipeline is detected through a
classification approach. We will explain each step in detail in the following sections.

@@

We reconstruct the tree of processes starting from the first process
created by the pipeline and identifying its child processes as the ones
that were created through \texttt{clone()} or \texttt{fork()}. We
create a process graph from this tree by adding edges corresponding to
file dependencies between processes. A file dependency is defined
between processes A and B if a file written by A is read by
B. Figure~\ref{fig:simple_script} shows an example of a process graph
constructed at this stage.

@@

bet ${input_image} ${bet_output} > bet_temp.out
echo "Voxels / Volume in brain mask:"
fslstats ${bet_output} -V
fslmaths ${bet_output} ${bet_output_binarized}
echo "Voxels / Volum in binarized brain mask:"
\rm bet_temp.out
\end{verbatim}
}
\end{minipage}
\captionof{figure}{Process graph constructed from an example brain extraction
    script. Every node in the graph is labeled using (1) a process id
    created by our reconstruction, (2) the name of the executable run
    by the process. Processes that read or write temporary files are
    represented with squares; other ones with circles. Plain edges
    represent the process tree (\texttt{fork()} or \texttt{clone()}
    system calls). Dashed edges represent file dependencies: temporary
    files are in yellow and result files are in green.} \label{fig:simple_script}
\medskip

@@

\paragraph{} In the next step, we classify the processes into three 
categories based on the process graph and the error matrix file 
mentioned in the previous Section:
\begin{enumerate}
\item Processes that read files that do not have errors and write files that do not 
have errors are \emph{transparent} (Figure~\ref{fig:processes}.a).
\item Processes that read files 
that do not have any error but write files that have errors 
\emph{create} errors in the pipeline (Figure~\ref{fig:processes}.b).
\item Processes that read files 
that have errors and write files that do not have errors \emph{remove} 
errors from the pipeline (Figure~\ref{fig:processes}.c).
\item Processes that read files that have errors and write files that also have errors are 
\emph{unknown} (Figure~\ref{fig:processes}.d).
\end{enumerate}

\begin{figure}[H]%\centering
    \begin{subfigure}{0.15\textwidth}
        \includegraphics[scale=0.35]{images/green.png}
        \caption{Transparent}
        \label{fig:green}
    \end{subfigure}
\hfill%\hfil
    \begin{subfigure}{0.15\textwidth}
    \includegraphics[scale=0.35]{images/red.png}
    \caption{Creates error}
    \label{fig:red}
\end{subfigure}
\hfill%\hfil
    \begin{subfigure}{0.15\textwidth}
    \includegraphics[scale=0.35]{images/blue.png}
    \caption{Removes error}
    \label{fig:blue}
\end{subfigure}
\hfill%\hfil
    \begin{subfigure}{0.15\textwidth}
    \includegraphics[scale=0.35]{images/yellow.png}
    \caption{Unknown}
    \label{fig:yellow}
\end{subfigure}
    \caption{Different type of classified process based on the input/output files.
  Dashed edges refer to the file dependencies between processes A and B if a file written bu A is read bu B.
  Solid black edges refer ti the relationship between parent and child processes.}
    \label{fig:processes}
\end{figure}

\paragraph{Challenge1:} The classification of processes that read or write temporary files is 
uncertain when these files are deleted during the execution. To address 
this issue, we replace every process P that writes temporary files with 
a modification of P that first calls P and then backs up all its output 
files to a read-only directory. This replacement is done by modifying 
the Unix PATH variable to point to a directory containing the modified
versions of the processes. This solution does not cover the temporary files
that are removed by P itself; this is not a problem since these files do not
play any role in the subsequent steps of the pipeline, by definition.

\paragraph{Challenge2:} Files written by multiple processes also lead to unknown classification 
labels. To address this issue for a file F written by processes in 
\textbf{P} = \{$P_{1}$, \ldots $P_{n}$\}, we (1) check that processes 
in \textbf{P} do not write concurrently to F, (2) we establish an order 
on \textbf{P} based on the creation timestamp of the processes, (3) we 
replace ever process $P_{i}$ in \textbf{P} by a wrapper that first 
calls $P_{i}$ and then backs up F to a read-only directory. Thus, 
multiple versions of F are saved and used in the analysis.

\paragraph{Challenge3:} Another challenge is to classify processes that read files that have
errors and write files that also have errors. In such situations, it
is not possible to determine from the pipeline results whether
the process created errors, removed some errors, or was
transparent. We label such processes \emph{unknown}. To address this issue, we developed an iterative approach that consists of the following steps:
\begin{enumerate}
  \item Run the pipeline in conditions 1 and 2; classify the
    processes as \emph{transparent}, \emph{create errors},
    \emph{remove errors} or \emph{unknown}.
  \item If there are \emph{unknown} processes then: replace each process P that create errors by a process Q that
    copies the results produced by process P in condition 1 to the pipeline output in condition 2.
  \item Repeat steps 1 and 2 until there is no \emph{unknown} process left.
\end{enumerate}

The replacement of process P at step 2 is done by replacing P
with a custom script produced by step 1. This custom script
copies the results obtained in condition 1 if it is invoked with the
arguments of a process that created errors, and calls P's original
executable otherwise. The replacement is done through the PATH variable, as before.
This algorithm converges to a process graph without any \emph{unknown} process after
a finite number of iterations.

\begin{figure}[H]
\centering
  \includegraphics[scale=0.5]{images/iterative_modif}
  \caption{iteration of proposed modification.}
  \label{fig:iterations}
\end{figure}

@@

In a previous study~\cite{Scaria2017}, we showed that
pre-processing pipelines of the Human Connectome
Project~\cite{Glasser2013} were sensitive to operating system
variation (see Figure \ref{fig:1}).
\begin{figure}[H]
%  \includegraphics{brain\_classification}
  \includegraphics[width=0.5\textwidth, height=6cm]{images/brain_classification.png}
  \caption{Tissue classification produced by the HCP pre-processing
    pipelines on subject 105216 (CentOS 6 vs CentOS 7).}
  \label{fig:1}
\end{figure}

\begin{figure}[H]
\centering
  \includegraphics[scale=0.8]{images/fnirt_result.png}
  \caption{Differences between FNIRT results (T2w\_acpc\_to\_MNI\_nonli.nii.gz), subject 104820 (centos6 vs. centos7).}
  \label{fig:fnirt_result}
\end{figure}

@@

Figure \ref{fig:summarized-graph} shows the annotated provenance graph of the
PreFreeSurfer pipeline executed on CentOS6 and CentOS7.
Each node in the graph represent an executed process in the pipeline.
Processes that created errors are shown in red, processes that removed errors
are in blue, and other processes are in green.  Squares denote
processes for which the classification is uncertain, due to temporary
files that were removed during the execution. Black edges link
sub-processes to their parents while dashed edges denote file
dependencies between processes (green edges: files with no errors; red
edges: files with errors; yellow edges: temporary files).

The processes that introduce errors in PreFreeSurfer are: linear
registration with “\emph{FLIRT}” (8 occurrences, in ACPC Alignment,
BrainExtraction, DistortionCorrection, AtlasRegistration), non-linear
registration with “\emph{FNIRT}” (3 occurrences, in BrainExtraction
and AltasRegistration), image warping with “\emph{new\_invwarp}” (3
occurrences, in BrainExtraction and AtlasRegistration).  In addition,
errors were observed in image mean and standard-deviation computations
with “\emph{fslstats}” (3 occurrences in BiasFieldCorrection), and in
masked image extrapolation with “\emph{fslmaths}” (1 occurrence, in
BiasFieldCorrection).  Besides, transformation format conversion with
“\emph{convertwarp}” (2 occurrences, in DistortionCorrection) was able
to remove errors.

@@

\begin{figure}[H]
\centering
  \includegraphics[scale=0.3]{images/pfs_chart.png}
  \caption{Number of occurrences of errors in PreFreeSurfer, red and blue bars 
indicate the processes which create and remove errors respectively.}
  \label{fig:pfs_chart}
\end{figure}

\begin{figure}[H]
\centering
  \includegraphics[scale=0.5]{images/pfs_table.png}
  \caption{Process name along with the pipeline elements that inbtroduce errors.}
  \label{fig:pfs_table}
\end{figure}

The processes that introduce errors in FreeSurfer are shown on Figure~\ref{fig:fs_error_table}.
Also, processes that remove errors in FreeSurfer represented on Figure~\ref{fig:fs_remove_table}.
\begin{figure}[H]
\centering
  \includegraphics[scale=0.5]{images/fs_error_table.png}
  \caption{Number of occurrences of errors in FreeSurfer pipeline, subject 103818, 
indicate the processes that create errors.}
  \label{fig:fs_error_table}
\end{figure}

\begin{figure}[H]
\centering
  \includegraphics[scale=0.5]{images/fs_remove_table.png}
  \caption{Number of occurrences of errors in FreeSurfer pipeline, subject 103818, 
indicate the processes that remove errors.}
  \label{fig:fs_remove_table}
\end{figure}

@@

pipeline amplify small numerical differences, they are numerically unstable.
Furthermore, math libraries evolve over time, leading to different numerical errors.
we listed some of the irreproducibility causes of the pipelines as we mentioned
in the previous section overally along with exact command arguments.

\begin{itemize}
\item FSLSTATS: Calculate statistics on an image (mean, standard deviation and etc.)
    \begin{lstlisting}[language=bash, basicstyle=\tiny]
STD=`${FSLDIR}/bin/fslstats $WD/T1wmulT2w_brain_norm_modulate.nii.gz -S` 
MEAN=`${FSLDIR}/bin/fslstats $WD/T1wmulT2w_brain_norm_modulate.nii.gz -M` 
    \end{lstlisting}
\item FSLMATH: Create and apply a binary mask to filling of holes to remove  any non-grey/white tissue
using a threshold created by FSLSTATS at Mean.
    \begin{lstlisting}[language=bash, basicstyle=\tiny]
{FSLDIR}/bin/fslmaths $WD/T1wmulT2w_brain_norm_modulate -thr $Lower -bin -ero -mul 255 
                      $WD/T1wmulT2w_brain_norm_modulate_mask ${CARET7DIR}/wb_command -volume-remove-islands 
                      $WD/T1wmulT2w_brain_norm_modulate_mask.nii.gz $WD/T1wmulT2w_brain_norm_modulate_mask.nii.gz
    \end{lstlisting}

\item FLIRT: Register cropped image to MNI152 (12 DOF)
    \begin{lstlisting}[language=bash, basicstyle=\tiny]
{FSLDIR}/bin/flirt -interp spline -in "$WD"/robustroi.nii.gz -ref "$Reference"
                   -omat "$WD"/roi2std.mat -out "$WD"/acpc_final.nii.gz
                   -searchrx -30 30 -searchry -30 30 -searchrz -30 30

{FSLDIR}/bin/flirt -interp spline -dof 6 -in ${WD}/Magnitude_warpped${TXw}
                    -ref ${TXwImage} -out ${WD}/Magnitude_warpped${TXw}2${TXwImageBasename}
                    -omat ${WD}/Fieldmap2${TXwImageBasename}.mat
                    -searchrx -30 30 -searchry -30 30 -searchrz -30 30
    \end{lstlisting}

\item FNIRT: non-linear registration to MNI
    \begin{lstlisting}[language=bash, basicstyle=\tiny]
{FSLDIR}/bin/fnirt --in=${T1wRestore} --ref=${Reference2mm} --aff=${WD}/xfms/acpc2MNILinear.mat
                   --refmask=${Reference2mmMask} --fout=${OutputTransform}
                   --jout=${WD}/xfms/NonlinearRegJacobians.nii.gz --refout=${WD}/xfms/IntensityModulatedT1.nii.gz
                   --iout=${WD}/xfms/2mmReg.nii.gz --logout=${WD}/xfms/NonlinearReg.txt
                   --intout=${WD}/xfms/NonlinearIntensities.nii.gz --cout=${WD}/xfms/NonlinearReg.nii.gz --config=${FNIRTConfig}

%~ Register to 2mm reference image (linear then non-linear)
%~ ${FSLDIR}/bin/flirt -interp spline -dof 12 -in "$Input" -ref "$Reference2mm" 
                    %~ -omat "$WD"/roughlin.mat -out "$WD"/"$BaseName"_to_MNI_roughlin.nii.gz -nosearch

%~ ${FSLDIR}/bin/fnirt --in="$Input" --ref="$Reference2mm" --aff="$WD"/roughlin.mat
                    %~ --refmask="$Reference2mmMask" --fout="$WD"/str2standard.nii.gz
                    %~ --jout="$WD"/NonlinearRegJacobians.nii.gz --refout="$WD"/IntensityModulatedT1.nii.gz
                    %~ --iout="$WD"/"$BaseName"_to_MNI_nonlin.nii.gz
                    %~ --logout="$WD"/NonlinearReg.txt --intout="$WD"/NonlinearIntensities.nii.gz
                    %~ --cout="$WD"/NonlinearReg.nii.gz --config="$FNIRTConfig"
    \end{lstlisting}

\item INVWARP \& APPLYWARP: Invert warp and transform dilated brain mask back into native space, and use it to mask input image
    \begin{lstlisting}[language=bash, basicstyle=\tiny]
${FSLDIR}/bin/invwarp --ref="$Reference2mm" -w "$WD"/str2standard.nii.gz -o "$WD"/standard2str.nii.gz
${FSLDIR}/bin/applywarp --rel --interp=nn --in="$ReferenceMask" --ref="$Input" -w "$WD"/standard2str.nii.gz -o "$OutputBrainMask"
    \end{lstlisting}

\item MRIS\_ANATOMICAL\_STATS: Measures a number of anatomical properties
    \begin{lstlisting}[language=bash, basicstyle=\tiny]
mris_anatomical_stats -mgz -cortex ../label/lh.cortex.label -f ../stats/lh.aparc.a2009s.stats -b -a ../label/lh.aparc.a2009s.annot -c ../label/aparc.annot.a2009s.ctab 0050299 lh white
    \end{lstlisting}

\item MRIS\_MAKE\_SURFACES: Generate pial surfaces, first pass, no T2 adjustment. Files created are Xh.pial, Xh.curv.pial, Xh.area.pial, Xh.thickness
    \begin{lstlisting}[language=bash, basicstyle=\tiny]
mris_make_surfaces -variablesigma ${VARIABLESIGMA} -white NOWRITE -aseg aseg.hires -orig white.deformed -filled filled.hires -wm wm.hires -sdir $SubjectDIR -mgz -T1 T1w_hires.norm "$SubjectID" lh
    \end{lstlisting}

\item MRI\_CONVERT:Normalize T1w and T2w images for the benefit of mris\_make\_surfaces

\#Resample the native space binary ROI file into "brainmask" space
    \begin{lstlisting}[language=bash, basicstyle=\tiny]
mri_convert -rl /subA/mri/T1w_hires.nii.gz -rt nearest /subA/mri/filled.mgz /subA/mri/filled.hires.mgz
    \end{lstlisting}

\item MRI\_NORMALIZE: Normalize the white-matter, optionally based on control points.
The input volume is converted into a new volume where white matter image values all range around 110.

\#Uses the norm volume, and creates the brain volume, making use of the aseg and masking with brainmask
    \begin{lstlisting}[language=bash, basicstyle=\tiny]
mri_normalize -aseg aseg.mgz -mask brainmask.mgz norm.mgz brain.mgz
    \end{lstlisting}

\item MRI\_APARC2ASEG:The volumes of the cortical labels will be different than
         reported by mris\_anatomical\_stats because partial volume information
         is lost when mapping the surface to the volume. The values reported by
         mris\_anatomical\_stats will be more accurate than the volumes from the
         aparc+aseg volume.

Label the cortex of a subject's segmented volume according to the edited surface labels
    \begin{lstlisting}[language=bash, basicstyle=\tiny]
mri_aparc2aseg --s SUBJ_NUM --volmask
\#slice labeling
mri_aparc2aseg --s ${subject} --labelwm --hypo-as-wm --rip-unknown --volmask --o ${subj_dir}/mri/wmdivided --annot aparc.split
    \end{lstlisting}

\item MRI\_EM\_REGISTER: creates a tranform in lta format
    \begin{lstlisting}[language=bash, basicstyle=\tiny]
mri_em_register -skull -t transforms/talairach.lta nu_noneck.mgz $FREESURFER_HOME/average/RB_all_withskull_2008-03-26.gca transforms/talairach_with_skull.lta
    \end{lstlisting}

\item MRIS\_SMOOTH: Smooths the tessellation of a cortical surface.
    \begin{lstlisting}[language=bash, basicstyle=\tiny]
mris_smooth -nw -seed 1234 ../surf/lh.orig.nofix ../surf/lh.smoothwm.nofix
    \end{lstlisting}

\item MRIS\_INFLATE: inflate cortical surface.
\#The command generates inflated and sulc files.
    \begin{lstlisting}[language=bash, basicstyle=\tiny]
mris_inflate -no-save-sulc surf/lh.smoothwm.nofix surf/lh.inflated.nofix
    \end{lstlisting}

\item MRIS\_CALC: mris\_calc' is a simple calculator that operates on
FreeSurfer curvatures and volumes. In most cases, the calculator functions
with three arguments: two inputs and an <ACTION> linking them.
Some actions, however, operate with only one input <file1>.
In all cases, the first input <file1> is the name of a FreeSurfer curvature overlay (e.g. rh.curv)
or volume file (e.g. orig.mgz). For two inputs, the calculator first assumes
that the second input is a file. If, however, this second input file doesn't exist,
the calculator assumes it refers to a float number, which is then processed according to <ACTION>.Note:
<file1> and <file2> should typically be generated on the same subject.
    \begin{lstlisting}[language=bash, basicstyle=\tiny]

mris_calc -o lh.area.mid lh.area.mid div2
    \end{lstlisting}

\item CONVERTWARP: Forward warp the fieldmap magnitude and register to TXw image (transform phase image as well).
    \begin{lstlisting}[language=bash, basicstyle=\tiny]
${FSLDIR}/bin/convertwarp --relout --rel --ref=${WD}/Magnitude --shiftmap=${WD}/FieldMap_ShiftMap${TXw}.nii.gz 
                          --shiftdir=${UnwarpDir} --out=${WD}/FieldMap_Warp${TXw}.nii.gz
    \end{lstlisting}
\item MRI\_CONVERT:

\item MRI\_SURF2SURF:

\item MRIS\_MAKE\_SURFACE:

\item MRI\_CA\_REGISTER:

\item MRIS\_CONVERT:

\end{itemize}

Mention that this was only possible because the unprocessed data was shared in the first place.
DICOM to Nifti conversion was out of scope and may introduce other issues.

@@

Our thechnique is able to characterize the stability of a pipeline's components automatically.
The numerical instability in the PreFreesurfer HCP pipeline arises mainly from
linear and non-linear registration processes implemented in FSL FLIRT and FNIRT.

There are a few ways to impede such instabilities:
\begin{itemize}
\item Use a single operating system
\item Containerize pipelines
\item Increase numerical precision
\item Be stricter on truncation and rounding standards (IEEE 754)
\item Build static executable
\end{itemize}

The results still suffer from small pertubations literally because of the fact that pipeline are not numerically stable.
The preferred solution is to detect and fix numerical instability of the pipeline instead of masking the problem.
These processes need to be reviewed to understand and correct the cause of instabilities. 
In this correction process, accuracy has to be considered in addition to stability.

\item In the future works we aim to measure the impact of such errors
on the specific domain of the neuroscience and how show the significant it is.

\item The purpose of the project is generalizing pipelines to get a reproducible execution 
on every operating system instead of choosing one main OS to.

@@

\begin{figure}[H]
  \includegraphics[width=\linewidth]{images/summarized_graph}
  \caption{An annotated summarized process graph from the PreFreesurfer pipeline.
Each node in the graph is represented in red, blue or green colors and is labeled 
using the name of the executable run by the interesting process. The close up nodes 
reveal each shrinked subgraph in details.}
  \label{fig:summarized-graph}
\end{figure}

\begin{figure}[H]
  \includegraphics[width=\linewidth]{images/graph}
  \caption{A complete process graph from the PreFreesurfer pipeline.
Full-resolution image available at \url{https://drive.google.com/open?id=174yyn8SuVOUcK5aRVw0bagjDanLD0FLt}.}
  \label{fig:complete-graph}
\end{figure}
\bibliographystyle{plain}
\bibliography{biblio}
